HPC--101\\Onboarding

### 2017-02-16 hpc_101/sections/02_cluster.tex
changed node description to table fixed up footnotes across multiple slides.  Also removed section about related services (matlab DCS)

\subsubsection{Hardware}
\begin{frame}
	\frametitle{{\craycs} -- Compute Nodes}
	\begin{itemize}
        \item All nodes are diskless with some RAM used for the Operating System
	\item 178 Standard nodes {\tiny(\emph{community})}
	  \begin{itemize}
            {\footnotesize
	    \item Two 10 core {\intel} Ivy-Bridge processors (\emph{20 cores total})
	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
            }
	  \end{itemize}          
	\item 6 Large memory nodes {\tiny(\emph{community})}
	  \begin{itemize}
            {\footnotesize
	    \item Four 10 core {\intel}  Ivy-Bridge processors (\emph{40 cores total})
	    \item 1TB of physical RAM with $\approx1000$GB of useable RAM
            }
	  \end{itemize}			
        \item 1 GPU node {\tiny(\emph{community})}
          \begin{itemize}
            {\footnotesize
	    \item Two 10 core {\intel} Haswell processors (\emph{20 cores total})
	    \item 128GB of physical RAM with $\approx118$GB of useable RAM
            \item 2 Nvidia Tesla K40 GPUs
            }
	  \end{itemize}
        \item 96 Owner nodes
          \begin{itemize}
            {\footnotesize
            \item 33 nodes have two 10 core {\intel} Haswell processors with 256 GB of RAM
            \item 62 nodes have two 12 core {\intel} Haswell processors with 128 GB of RAM
            \item 1 GPU node
            }
          \end{itemize}
	\item CentOS Linux
	\end{itemize}
\end{frame}

>>>>

\subsubsection{Hardware}
\begin{frame}
  \frametitle{{\craycs} -- Compute Nodes}
  \begin{table}
    \centering
    \resizebox{\textwidth}{!}{%
      \begin{tabular}{|l|l|l|l|l|l|l|l|}
        \multicolumn{8}{c}{ {\large {\textbf{ Community Nodes\footnote{\label{node_note}\tiny Nodes are diskless and run CentOS linux}} } } }\\ \hline
        \textbf{Type} & \textbf{Cores} & \textbf{Ivy-Bridge} & \textbf{Haswell} & \textbf{Broadwell} &
        \textbf{128 GB}\footnote{\label{mem_limit}\tiny Usable RAM: $\approx118$ GB of 128 GB, $\approx246$ GB of 256 GB, $\approx1003$ GB of 1 TB} & \textbf{256 GB}\textsuperscript{\ref{mem_limit}}  & \textbf{1 TB}\textsuperscript{\ref{mem_limit}}  \\ \hline
        Standard & 20 & 178 &  &  & 178 &  &  \\ \hline
        Large & 40 & 6 &  &  &  &  & 6 \\ \hline
        GPU\footnote{\label{gpu_v1}\tiny 2 x Nvidia Tesla K40} & 20 &  & 1 &  & 1 &  & \\ \hline
      \end{tabular}%
    }
  \end{table}

  \begin{table}
    \centering
    \resizebox{\textwidth}{!}{%
      \begin{tabular}{|l|l|l|l|l|l|l|l|}
        \multicolumn{8}{c}{ {\large {\textbf{ Condo Nodes}\textsuperscript{\ref{node_note}} } }} \\ \hline
        \textbf{Type} & \textbf{Cores} & \textbf{Ivy-Bridge} & \textbf{Haswell} & \textbf{Broadwell} & \textbf{128 GB}\textsuperscript{\ref{mem_limit}} & \textbf{256 GB}\textsuperscript{\ref{mem_limit}} & \textbf{1 TB}\textsuperscript{\ref{mem_limit}} \\ \hline
        Standard & 24 & & 62 & & 62 & & \\ \hline
        Standard & 20 & & & (5)\footnote{\label{future}\tiny April 2017 additions} & (5)\textsuperscript{\ref{future}} & & \\ \hline
        Custom & 20 &  & 33 & (4)\textsuperscript{\ref{future}} &  & 33 & (4)\textsuperscript{\ref{future}} \\ \hline
        GPU\textsuperscript{\ref{gpu_v1}}     & 20 &  & 1 &  & 1 &  &  \\ \hline
      \end{tabular}%
    }
    \end{table}
  \end{frame}

### 2017-02-16 hpc_101/sections/04_cluster_interaction.tex
changed node description to table fixed up footnotes across multiple slides.  Also removed section about related services (matlab DCS)

\subsection{Software}
\begin{frame}
	\frametitle{Modules}
	\begin{block}{Modules}\tiny
	A tool to help users manage their Unix or Linux shell environment, by allowing groups of related environment-variable settings to be made or removed dynamically.\footnotemark

	\end{block}
	\begin{itemize}
		\item We globally install frequently requested software packages and create modules for all users to access
		\item Access to modules is via the \textbf{module} command
		\begin{itemize}\footnotesize
			\item `module avail' -- list installed modules
			\item `module show $<$module name$>$' -- Show what actions a module performs
			\item `module load $<$module name$>$' -- Loads the named module
			\item `module purge' -- Unload all loaded modules
		\end{itemize}
		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
	\end{itemize}
	\footnotetext[1]{\tiny \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
\end{frame}

@@

\begin{frame}
\frametitle{Compilers}
	\begin{itemize}
		\item We have the {\intel}, GNU (gcc, g$++$), {\cray} \& PGI{\regtrademark} compilers
		\item Compiling must take place on a compute node
		\begin{itemize}
			\item Interactive sessions are useful for compiling software
			\item Sandbox nodes mirror the environment the compute nodes provide and are ideal for compilation
			\item Login nodes \textbf{do not} load all the software and libraries found on the compute nodes
		\end{itemize}
		\item {\intel} compilers are recommended for best performance
		\begin{itemize}
			\item {\intel} 2013 compilers:
			\begin{itemize}
				\item module load intel/ics -- Loads {\intel} compilers: {\tiny \textbf{icc}, \textbf{ifort}, \textbf{icpc}}
				\item module load intel/impi -- Loads {\intel} MPI wrapper: {\tiny\textbf{mpiicc}, \textbf{mpiifort}, \textbf{mpiicpc}}
			\end{itemize}
			\item {\intel} 2016 compilers:
			\begin{itemize}
				\item We have 2 floating seats for {\intel} 2016 compiler
				\item intel\_2016/ics 
				\item intel\_2016/impi 
			\end{itemize}
		\end{itemize}			
	\end{itemize}
\end{frame}

@@

\subsection{SLURM Partitions}
\begin{frame}
\footnotesize
\frametitle{Partitions}
\begin{block}{What is a partition?}\footnotesize
A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.  Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc. Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnotemark
\end{block}
\begin{itemize}
	\item The {\craycs} currently has seven public partitions:~\\\textbf{community.q}, \textbf{exclusive.q}, \textbf{lm.q}, \textbf{gpu.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
\end{itemize}
\footnotetext[2]{\tiny \url{http://slurm.schedmd.com/quickstart.html}}
\end{frame}

@@

\subsection{Interactive Job -- srun}
\frametitle{SLURM Job Scripts}
\begin{frame}
\frametitle{Interactive Job with SLURM}
	\begin{block}{Interactive session}\tiny
	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00 \ddash{}pty /bin/bash
          ~\\
          ~\\
          \hrule
          ~\\
	$[$login \ctilde$]$\$ srun -I -p sb.q -N 1 -c 1 -n 1 -t 0-01:00:00 \ddash{}pty /bin/bash

>>>>

\subsection{Software}
\begin{frame}
	\frametitle{Modules}
	\begin{block}{Modules}\tiny
	  A tool to help users manage their Unix or Linux shell environment, by allowing groups of related environment-variable settings to be made or removed dynamically.\footnote{\label{wiki_module}\tiny\
             \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}

	\end{block}
	\begin{itemize}
		\item We globally install frequently requested software packages and create modules for all users to access
		\item Access to modules is via the \textbf{module} command
		\begin{itemize}\footnotesize
			\item `module avail' -- list installed modules
			\item `module show $<$module name$>$' -- Show what actions a module performs
			\item `module load $<$module name$>$' -- Loads the named module
			\item `module purge' -- Unload all loaded modules
		\end{itemize}
		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
	\end{itemize}

@@

\begin{frame}
\frametitle{Compilers}
	\begin{itemize}
		\item We have the {\intel} \& GNU (gcc, g$++$) compilers
		\item Compiling must take place on a compute node
		\begin{itemize}
			\item Interactive sessions are useful for compiling software
			\item Sandbox nodes mirror the environment the compute nodes provide and are ideal for compilation
			\item Login nodes \textbf{do not} load all the software and libraries found on the compute nodes
		\end{itemize}
		\item {\intel} compilers are recommended for best performance
		\begin{itemize}
			\item {\intel} 2013 compilers:
			\begin{itemize}
				\item module load intel/ics -- Loads {\intel} compilers: {\tiny \textbf{icc}, \textbf{ifort}, \textbf{icpc}}
				\item module load intel/impi -- Loads {\intel} MPI wrapper: {\tiny\textbf{mpiicc}, \textbf{mpiifort}, \textbf{mpiicpc}}
			\end{itemize}
			\item {\intel} 2016 compilers:
			\begin{itemize}
				\item We have 2 floating seats for {\intel} 2016 compiler
				\item intel\_2016/ics 
				\item intel\_2016/impi 
			\end{itemize}
		\end{itemize}			
	\end{itemize}
\end{frame}

@@

\subsection{SLURM Partitions}
\begin{frame}
\footnotesize
\frametitle{Partitions}
\begin{block}{What is a partition?}\footnotesize
A partition can be thought of as a group of nodes/resources divided into possibly overlapping sets.  Each partition can be considered as a job queue, each of which has an assortment of constraints such as job size limit, job time limit, users permitted to use it, etc. Priority-ordered jobs are allocated nodes within a partition until the resources (nodes, processors, memory, etc.) within that partition are exhausted.\footnote{\label{schedmd_qstart}\tiny \url{http://slurm.schedmd.com/quickstart.html}}
\end{block}
\begin{itemize}
	\item The {\craycs} currently has seven public partitions:~\\\textbf{community.q}, \textbf{exclusive.q}, \textbf{lm.q}, \textbf{gpu.q}, \textbf{sb.q}, \textbf{kill.q}, \textbf{htc.q}
	\item Jobs submitted to kill.q and htc.q can be preempted by jobs in other partitions
\end{itemize}
\end{frame}

@@

\subsection{Interactive Job -- srun}
\frametitle{SLURM Job Scripts}
\begin{frame}
  \frametitle{Interactive Job with SLURM}
  \begin{block}{Interactive session}\tiny
    $[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 1 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00:00 \ddash{}pty /bin/bash
    ~\\
    ~\\
    ~\\
    \hrule\begin{semiverbatim}$[$login \ctilde$]$\$ srun -I -p sb.q -N 1 -c 1 -n 1 -t 0-01:00:00 \ddash{}pty /bin/bash\end{semiverbatim}
